Bayesian RKHS-based methods in functional regression\support{Support information of the article.}

### 2022-04-06 paper/paper.tex
More corrections

\startlocaldefs
\numberwithin{equation}{section}
\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{prop}{Proposition}[section]

@@

%% Environment for colored omments
\newenvironment{comment}[1][comment-red]
{
\noindent \color{#1}
}
{
\color{black}
}

@@

\begin{comment}
Posibles temas de desarrollo teórico:
  \begin{itemize}
    \item Enfoque de procesos Gaussianos como priors en modelos RKHS, eligiendo un kernel \(K'\) que tenga buenas propiedades con respecto a \(K\). Distribuciones conjugadas.
    \item Temas de consistencia de la posteriori: concentración hacia el ``verdadero'' valor el parámetro o bien convergencia hacia la ``verdadera'' posteriori.
    \item Distancia máxima entre las funciones de \(\Hcal_{0,p}(K)\) y las de \(\Hcal_0(K)\) (ó \(\Hcal(K)\)).
    \item Robustez de las distribuciones ante pequeñas perturbaciones.
    \item  En \citet[Sec.~2.5]{grollemund2019bayesian}, los autores proponen un estimador Bayesiano para sus funciones ``escalón'' \(\beta_\theta(\cdot)\) utilizando una pérdida cuadrática, de la forma
  \[
    \hat \beta \in \arg\min_{d\in L^2[0, 1]} \int\int (\beta_\theta(t) - d(t))^2 \pi(\theta\mid \mathcal D_n)\, dt\, d\theta.
  \]
  ¿Podríamos hacer algo parecido con nuestro parámetro funcional \(\alpha \in \Hcal_{0,p}(K)\)?
  \end{itemize}
\end{comment}

@@

Lastly, a particularly useful approach in statistics is to regard \(\Hcal(K)\) as an isometric copy of a well-known space. Specifically, via \textit{Loève's isometry} one can establish a congruence between \(\Hcal(K)\) and the linear span of the process, \(\mathcal L(X)\), in the space of all random variables with finite second moment, \(L^2(\Omega)\) (see Lemma 1.1 in \citet{lukic2001stochastic}). This isometry is essentially the completion of the correspondence
  \[
  \sum_{i=1}^p a_i X(t_i) \longleftrightarrow \sum_{i=1}^p a_i K(t_i, \cdot),
\]
and can be formally defined as \(\Psi^{-1}_X(U)(t) = \E[U X(t)]\) for \(U \in \mathcal L(X)\).
Despite the close connection between the process \(X\) and the space \(\Hcal(K)\), special care must be taken when dealing with concrete realizations of the process, since under rather general conditions the trajectories of \(X\) do not belong to the corresponding RKHS with probability one \citep[see for example][Cor.~7.1]{lukic2001stochastic}. As a consequence, the expression \(\dotprod{x}{f}_K\) is ill-defined and lacks meaning when \(x\) is a realization of \(X\). However, following Parzen's approach in his seminal work on time series analysis \citep[Th.~4E]{parzen1961approach}, we can leverage Loève's isometry and identify \(\dotprod{x}{f}_K \) with \( \Psi_x(f) := \Psi_X(f)(\omega)\), for \(x=X(\omega)\) and \(f\in \Hcal(K)\). This notation often proves to be useful and convenient.

@@

The general idea will be to impose a prior distribution on these free parameters (and possibly others) to eventually derive a posterior model after incorporating the available sample information.  Moreover, as we said before, with a slight abuse of notation we will understand the expression \(\dotprod{x}{\alpha}_K\) as \(\Psi_x(\alpha)\), where \(\Psi\) is Loève's isometry. Hence, taking into account that \(\Psi_X(K(t, \cdot)) = X(t)\) by definition, we have
\[
\dotprod{x}{\alpha}_K \equiv \sum_{j=1}^p \beta_j x(t_j).
\]

In view of~\eqref{eq:h0p}, to set a prior distribution on the unknown function \(\alpha\) (that is, a prior distribution on the functional space \(\Hcal_{0,p}(K)\)) it suffices to consider \(p\)-dimensional continuous prior distributions for the coefficients \(\beta_j\) and the times \(t_j\) separately. Thanks to this parametric approach, the challenging task of setting a prior distribution on a space of functions is considerably simplified, while simultaneously not constraining the model to any specific distribution (in contrast to Gaussian process regression methods). Moreover, note that starting from a probability distribution \(\mathbb{P}_0\) on \(\Hcal_0(K)\) we can obtain a probability distribution \(\mathbb{P}\) on \(\Hcal(K)\) simply by defining \(\mathbb{P}(B) = \mathbb{P}_0(B\cap \Hcal_0(K))\) for all Borel sets \(B\). Consequently, our simplifying assumption on \(\alpha\) is not very restrictive, since any prior distribution on \(\Hcal_0(K)\) can be directly extended to a prior distribution on \(\Hcal(K)\). Even though we actually work on \(\Hcal_{0,p}\), the parameter \(p\) is fundamentally...

In any case, after selecting a suitable prior distribution \(\pi(\theta)\) for the parameter vector \(\theta\) on the chosen finite-dimensional parameter space, we can resort to Bayes' theorem to perform the inference step, which in the case of i.i.d. samples amounts to
\begin{equation}\label{eq:bayes-theorem}
  \pi(\theta \mid \mathcal D_n) \propto \left[ \prod_{i=1}^n \pi(Y_i\mid X_i, \theta) \right]\pi(\theta).
\end{equation}
In Sections~\ref{sec:rkhs-linear-model} and~\ref{sec:rkhs-logistic-model} we proceed to specify the parameter spaces, prior distributions and concrete models for \(\pi(Y\mid X,\theta)\) considered in the case of functional linear regression and functional logistic regression, respectively.

\incomment{Disminuir mucho el párrafo de MCMC, simplemente comentarlo de pasada y en todo caso alguna referencia a una survey.}

A last technicality to discuss is the fact that in~\eqref{eq:bayes-theorem} we have omitted the possibly intractable integral related to the normalizing constant. Nevertheless, we can sample from the posterior distribution using an algorithm of the family of Markov Chain Monte Carlo (MCMC) sampling methods. These are subset of the more general Monte Carlo simulation methods, which are specifically designed to sample from a multi-dimensional distribution when only a function proportional to its density (or its log-density) is known. Generally speaking, these algorithms obtain samples from a continuous random variable by generating Markov chains whose stationary distribution is proportional to the given function. In short, they operate by building stochastic processes of random walkers that travel the parameter space, selecting where to move based on certain criteria that make them more likely to generate samples from the intended distribution. The Metropolis-Hastings algorithm \citep[e.g.][]{chib1995understanding} is arguably the most well-known method in this regard.

@@

In the case of functional linear regression, the RKHS model considered (with a functional parameter \(\alpha\in \Hcal_{0,p}(K)\)) is
\begin{equation}\label{eq:rkhs-model-linear}
  Y = \alpha_0 + \Psi_X(\alpha) + \epsilon = \alpha_0 + \sum_{j=1}^p \beta_j X(t_j) + \epsilon,
\end{equation}
where \(\beta_j \in \R\), \(t_j \in [0, 1]\), \(\alpha_0\in\R\), and \(\epsilon\) is an independent normally-distributed error term with \(\E[\epsilon]=0\) and \(\Var(\epsilon)=\sigma^2\in \R^+\). When \(p\) is fixed, the parameter space of dimension \(2p + 2\) becomes \(\Theta_p = \R^p \times [0, 1]^p \times \R \times \R^+\), and a generic element will be denoted by \(\theta = (\beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2) \equiv (b, \tau, \alpha_0, \sigma^2)\). Before proceeding any further, observe that we can rewrite model~\eqref{eq:rkhs-model-linear} in a more explicit and practical fashion in terms of the available sample information in \(\mathcal D_n\). For \(\theta \in \Theta_p\), the reinterpreted model assumes the form
\begin{equation}\label{eq:rkhs-model-linear-2}
  Y_i \mid X_i, \theta \ \stackrel{\text{i.i.d.}}{\sim} \mathcal N\left(\alpha_0 + \sum_{j=1}^p \beta_j X_i(t_j), \ \sigma^2\right), \quad i =1,\dots, n.
\end{equation}

It is worth mentioning that the model remains linear in the sense that it fundamentally involves a random variable belonging to the linear span of the process \(X\) in \(L^2(\Omega)\). Also, note that given the time instants \(t_j\) the model becomes a multiple linear model with the \(X(t_j)\) as scalar covariates. As a matter of fact, this RKHS model is particularly suited as a basis for variable selection methods, and also entails the classical \(L^2\)-model~\eqref{eq:l2-linear-model} under certain conditions \citep[see][Sec.~3]{berrendero2020general}. In addition, as far as generalizations go the model is easily extensible, seeing as how several functional covariates belonging to the RKHS could be added. In that case, as argued in \citet{grollemund2019bayesian} for a similar situation, if we were to set a prior distribution on all the parameters involved, we could recover the full posterior by looking alternately at the posterior distribution of each covariate conditional on the rest of them.

The prior distribution \(\pi(\theta)\) suggested for the parameter \(\theta \in \Theta_p\) is given by
\begin{align}\label{eq:prior-linear}
  \begin{split}
  \pi(\alpha_0, \sigma^2)              & \propto 1/\sigma^2,                                                     \\
  \tau                     & \sim \mathcal U([0, 1]^p),                                              \\
  \beta\mid \tau, \sigma^2 & \sim \mathcal N_p(b_0, g\sigma^2(G_\tau + \eta I)^{-1}),
\end{split}
\end{align}
where \(I\) is the identity matrix, \(G_\tau = \mathcal X_\tau'\mathcal X_\tau\) is the Gram matrix corresponding to the matrix \(\mathcal X_\tau = (X_i(t_j))_{i,j}\), and \(b_0\in \R^p, \ g \in \R\) and \(\eta \in \R^+\) are hyperparameters of the model. On the one hand, note the use of a joint prior distribution on \(\alpha_0\) and \(\sigma^2\), which is a widely used non-informative prior known as Jeffrey's prior \citep{jeffreys1946invariant}. In any event, the estimation of \(\alpha_0\) is straightforward, since \(\alpha_0 = \E[Y]\) by~\eqref{eq:rkhs-model-linear-2}, so we could have left this parameter out of the model altogether. On the other hand, the prior on \(\beta\) is a slight modification of the well-known Zellner's g-prior \citep{zellner1986assessing} in which a regularizing term is added to avoid ill-conditioning problems in the Gram matrix, obtaining a ridge-like Zellner prior controlled by the tuning parameter \(\eta\) \citep{baragatti2012study}. \incomment{Comentar la distribución a priori en su forma completa, decir que son independientes.}

\begin{comment}
Pequeña proposición con la cuenta de la posteriori
\[
\pi(\beta, \tau, \alpha_0, \log\sigma\mid \mathcal D_n) \propto \frac{|G_\tau|^{1/2}}{\sigma^{p+n}} \exp\left\{ -\frac{1}{2\sigma^2} \left(\|Y- \alpha_0\boldsymbol{1} - \mathcal X_\tau\beta\|^2 + \frac{1}{g}(\beta - b_0)'G_\tau(\beta - b_0) \right) \right\}.
\]

When performing the empirical posterior approximation, on each of the \(M\) steps of the iterative MCMC algorithm we get an approximate sample \(\theta^{(m)*}\) of the posterior distribution \(\pi(\theta\mid \mathcal D_n)\). Assuming now a previously unseen test set \(\mathcal D'_{n'}\) in the same conditions as \(\mathcal D_n\), we propose to construct three different kinds of estimators based on the MCMC samples: \incomment{Recordemos que el algoritmo nos da una observación del vector de parámetros,y hay varias estrategias para usarlas.}

\begin{comment}
  Comentar que una es promediar y luego predecir, y la otra es al revés. Separar un poco lo de selección de variables, comentando que nos concentramos solo en los tiempos y descartamos el resto (se consideran nuisance) (promediar los subvectores de tiempos y luego predecir como en la estrategia A, pero usando modelo multivariante). Cambiar los nombres.
\end{comment}

\begin{description}
  \item[Point estimates] If we consider a point-estimate statistic \(T\) to act as a summary of the posterior distribution, we can get the corresponding estimates \(\hat{\theta}=(\hat \beta, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T(\{\theta^{(m)*}\})\) and predict the responses in the usual way \incomment{Following model...} by
  \[
    \hat Y_i = \hat{\alpha}_0 + \sum_{j=1}^p \hat{\beta}_j X_i(\hat{t}_j), \quad i=1,\dots, n'.
  \]
  The choice of a specific summary statistic results in a veiled assumption of an underlying loss function between the estimated and real parameters. In general, the mean is more sensitive to outliers while the median is more robust, but the latter assumes a \(L^1\)-type loss function. On the other hand, the mode is also a good candidate because it represents the point of highest probability. At any rate, these decisions are strongly dependent on several factors such as the skewness or the number of modes in the resulting posterior distribution, and thus should be made on a case-by-case basis.
  \item[Distributional estimates] We can also look at the approximate posterior distribution as a whole and compute the distribution of the ``sampled responses'' at each step of the chain following model~\eqref{eq:rkhs-model-linear-2}:
  \[
  Y^{(m)*} := \left\{Y_i^{(m)*} \equiv Y_i \mid X_i, \theta^{(m)*}:\ i=1,\dots,n'\right\}, \quad m=1,\dots,M.
  \]
  Then, we can take the mean of all such samples as a proxy for each response variable:
  \[
  \hat Y_i = \frac{1}{M}\sum_{m=1}^M Y_i^{(m)*}, \quad i=1,\dots,n'.
  \]
  \item[Variable selection estimates] Lastly, we can focus only on the marginal posterior distribution of \(\tau\mid\mathcal D_n\) and perform variable selection via a point-estimate statistic \(T\). Specifically, we can consider the times \(\hat t_j = T(\{t_j^{(m)*}\})\) and reduce the original data set to the scalar variables \(\{X_i(\hat t_j):\ i=1, \dots,n, \ j=1,\dots,p\}\). After this selection, we can treat the problem as a finite-dimensional regression task and use any of the well-known multivariate algorithms suited for this situation.
\end{description}

\begin{comment}
Comentar después en la sección de resultados cómo se comparan, cuál es más estable.....
\end{comment}

@@

In the case of functional logistic regression, we can view the binary response variable \(Y\in\{0, 1\}\) as a Bernoulli random variable given the regressor \(X=x \in L^2[0, 1]\), and suppose that \(\log\left(p(x)/(1-p(x))\right)\) is linear in \(x\), where \(p(x)=\mathbb P(Y=1\mid X=x)\). Then, a RKHS approach given by the following functional logistic equation:
\begin{equation}\label{eq:rkhs-model-logistic}
  \mathbb P(Y=1 \mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \Psi_X(\alpha)\}}, \quad \alpha \in \Hcal_{0,p}(K).
\end{equation}
Following our previous strategy, after incorporating the sample information we can rewrite the classification model~\eqref{eq:rkhs-model-logistic} as
\begin{equation}\label{eq:rkhs-model-logistic-2}
Y_i \mid X_i,\theta \ \stackrel{\text{i.i.d.}}{\sim} \operatorname{Bernoulli}(p_i), \quad i=1,\dots, n,
\end{equation}
with
\begin{equation}\label{eq:rkhs-model-logistic-2-parameter}
  p_i = \mathbb P(Y_i=1 \mid X_i,\theta) = \frac{1}{\displaystyle 1 + \exp\left\{-\alpha_0 - \sum_{j=1}^p \beta_j X_i(t_j)\right\}}, \quad i=1,\dots, n,
\end{equation}
where in turn \(\beta_j\in\R\), \(t_j\in[0, 1]\) and \(\alpha_0\in\R\).

@@

As far as prior distributions go, we propose to use the same ones as we did in~\eqref{eq:prior-linear} for the regression approach \incomment{Me gustaría probar con otras distribuciones a priori, especialmente para los coeficientes \(\beta\), ya que aquí no sé si tiene tanto sentido mantener \(\sigma^2\) como parámetro y usar el prior de Zellner. ¿Sabéis de alguna que se use en el caso de regresión logística?}. However, in this case the parameter \(\sigma^2\) is a nuisance parameter and can be safely ignored at the end, since it does not appear in the final model~\eqref{eq:rkhs-model-logistic-2}.

\begin{comment}
  Misma pregunta que antes: ¿ponemos la expresión de la posteriori?
  \begin{align*}
\log \pi(\beta, \tau, \alpha_0, \log\sigma\mid \mathcal D_n) & \propto \sum_{i=1}^n \left[ \left(\alpha_0 + \Psi_{x_i}(\alpha)\right)y_i - \log\left(1 + \exp\left\{\alpha_0 + \Psi_{x_i}(\alpha)\right\}\right)\right]\\
\quad &+ \frac{1}{2}\log |G_\tau| - p\log \sigma -\frac{1}{2g\sigma^2} (\beta - b_0)'G_\tau(\beta - b_0).
\end{align*}

@@

Lastly, the variable selection method is essentially the same as in the case of linear regression: we select \(p\) times from each trajectory based on the posterior distribution of \(\tau\mid \mathcal D_n\), and then feed the reduced dataset to a multivariate binary classification procedure.

@@

One of the key decisions in our Bayesian modeling scheme was whether to consider the number of components \(p\) as a member or the parameter space and integrate it into the model. While theoretically we could impose a prior distribution on \(p\) as well, we found that it would have some unwanted practical implications. For instance, it would make the implementation more complex, since the dimensionality of the parameters \(\beta\) and \(\tau\) would need to be fixed at a certain maximum value, but in the MCMC algorithm the working value of \(p\) would vary from one iteration to the next. In this case we would face a label switching issue \citep[c.f.][Sec.~2.3]{grollemund2019bayesian}, as we would have no way of knowing which set of parameters is ``active'' at any given time. Another possibility would be to adapt a purely Bayesian model selection technique \citep[see][Ch.~7]{piironen2017comparison, gelman1995bayesian} to our framework, or even derive some model aggregation methods to combine the posterior distributions obtained for different-sized models. In the end, for the sake of simplicity we decided to let \(p\) be an hyperparameter, so that we could use any model selection criteria (e.g. BIC, DIC, \ldots) to select its optimal value \incomment{deberíamos elegir uno de los criterios?}.

>>>>

\startlocaldefs
\numberwithin{equation}{section}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}{Proposition}[section]

@@

%% Environment for colored comments
\newenvironment{comment}[1][comment-red]
{
\noindent \color{#1}
}
{
\color{black}
}

@@

\begin{comment}
Posibles temas de desarrollo teórico:
  \begin{itemize}
    \item Enfoque de procesos Gaussianos como priors en modelos RKHS, eligiendo un kernel \(K'\) que tenga buenas propiedades con respecto a \(K\). Distribuciones conjugadas.
    \item Temas de consistencia de la posteriori: concentración hacia el ``verdadero'' valor el parámetro o bien convergencia hacia la ``verdadera'' posteriori.
    \item Distancia máxima entre las funciones de \(\Hcal_{0,p}(K)\) y las de \(\Hcal_0(K)\) (ó \(\Hcal(K)\)).
    \item Robustez de las distribuciones ante pequeñas perturbaciones.
    \item  En \citet[Sec.~2.5]{grollemund2019bayesian}, los autores proponen un estimador Bayesiano para sus funciones ``escalón'' \(\beta_\theta(\cdot)\) utilizando una pérdida cuadrática, de la forma
  \[
    \hat \beta \in \arg\min_{d\in L^2[0, 1]} \int\int (\beta_\theta(t) - d(t))^2 \pi(\theta\mid \mathcal D_n)\, dt\, d\theta.
  \]
  ¿Podríamos hacer algo parecido con nuestro parámetro funcional \(\alpha \in \Hcal_{0,p}(K)\)?
  \item Considerar oras distribuciones a priori (especialmente en regresión logística).
  \end{itemize}
\end{comment}

@@

Lastly, a particularly useful approach in statistics is to regard \(\Hcal(K)\) as an isometric copy of a well-known space. Specifically, via \textit{Loève's isometry} one can establish a congruence between \(\Hcal(K)\) and the linear span of the process, \(\mathcal L(X)\), in the space of all random variables with finite second moment, \(L^2(\Omega)\) \citep[see Lemma 1.1 in][]{lukic2001stochastic}. This isometry is essentially the completion of the correspondence
  \[
  \sum_{i=1}^p a_i X(t_i) \longleftrightarrow \sum_{i=1}^p a_i K(t_i, \cdot),
\]
and can be formally defined as \(\Psi^{-1}_X(U)(t) = \E[U X(t)]\) for \(U \in \mathcal L(X)\).
Despite the close connection between the process \(X\) and the space \(\Hcal(K)\), special care must be taken when dealing with concrete realizations of the process, since under rather general conditions the trajectories of \(X\) do not belong to the corresponding RKHS with probability one \citep[see for example][Cor.~7.1]{lukic2001stochastic}. As a consequence, the expression \(\dotprod{x}{f}_K\) is ill-defined and lacks meaning when \(x\) is a realization of \(X\). However, following Parzen's approach in his seminal work on time series analysis \citep[Th.~4E]{parzen1961approach}, we can leverage Loève's isometry and identify \(\dotprod{x}{f}_K \) with \( \Psi_x(f) := \Psi_X(f)(\omega)\), for \(x=X(\omega)\) and \(f\in \Hcal(K)\). This notation often proves to be useful and convenient.

@@

The general idea will be to impose a prior distribution on these free parameters (and possibly others) to eventually derive a posterior model after incorporating the available sample information.  Moreover, as we said before, with a slight abuse of notation we will understand the expression \(\dotprod{x}{\alpha}_K\) as \(\Psi_x(\alpha)\), where \(\Psi\) is Loève's isometry. Hence, taking into account that \(\Psi_X(K(t, \cdot)) = X(t)\) by definition, we can write \(\dotprod{x}{\alpha}_K \equiv \sum_j \beta_j x(t_j)\).

In view of~\eqref{eq:h0p}, to set a prior distribution on the unknown function \(\alpha\) (that is, a prior distribution on the functional space \(\Hcal_{0,p}(K)\)) it suffices to consider \(p\)-dimensional continuous prior distributions for the coefficients \(\beta_j\) and the times \(t_j\) separately. Thanks to this parametric approach, the challenging task of setting a prior distribution on a space of functions is considerably simplified, while simultaneously not constraining the model to any specific distribution (in contrast to, for instance, Gaussian process regression methods). Moreover, note that starting from a probability distribution \(\mathbb{P}_0\) on \(\Hcal_0(K)\) we can obtain a probability distribution \(\mathbb{P}\) on \(\Hcal(K)\) simply by defining \(\mathbb{P}(B) = \mathbb{P}_0(B\cap \Hcal_0(K))\) for all Borel sets \(B\). Consequently, our simplifying assumption on \(\alpha\) is not very restrictive, since any prior distribution on \(\Hcal_0(K)\) can be directly extended to a prior distribution on \(\Hcal(K)\). Even though we actually work on \(\Hcal_{0,p}(K)\), the parameter \(p\) is fundamentally...\incomment{completar}

In any case, after selecting a suitable prior distribution \(\pi(\theta)\) for the finite-dimensional parameter vector \(\theta\), we can resort to Bayes' theorem to perform the inference step, which in the case of i.i.d. samples amounts to
\begin{equation}\label{eq:bayes-theorem}
  \pi(\theta \mid \mathcal D_n) \propto \left[ \prod_{i=1}^n \pi(Y_i\mid X_i, \theta) \right]\pi(\theta).
\end{equation}
In Sections~\ref{sec:rkhs-linear-model} and~\ref{sec:rkhs-logistic-model} we proceed to specify the parameter spaces, prior distributions and concrete models for \(\pi(Y\mid X,\theta)\) considered in the case of functional linear regression and functional logistic regression, respectively. Even though in~\eqref{eq:bayes-theorem} we have omitted the possibly intractable integral related to the normalizing constant, sampling from the (approximate) posterior distribution can be accomplished via Markov Chain Monte Carlo (MCMC) methods \citep[e.g.][]{brooks2011handbook}.

@@

In the case of functional linear regression, the RKHS model considered (with parameter \(\alpha\in \Hcal_{0,p}(K)\)) is given by
\begin{equation}\label{eq:rkhs-model-linear}
  Y = \alpha_0 + \Psi_X(\alpha) + \epsilon = \alpha_0 + \sum_{j=1}^p \beta_j X(t_j) + \epsilon,
\end{equation}
where \(\beta_j \in \R\), \(t_j \in [0, 1]\), \(\alpha_0\in\R\), and \(\epsilon\) is an independent normally-distributed error term satisfying \(\E[\epsilon]=0\) and \(\Var(\epsilon)=\sigma^2\in \R^+\). When \(p\) is fixed, the parameter space of dimension \(2p + 2\) becomes \(\Theta_p = \R^p \times [0, 1]^p \times \R \times \R^+\), and in the sequel a generic element of this space will be denoted by \(\theta = (\beta_1,\dots, \beta_p, t_1,\dots, t_p, \alpha_0, \sigma^2) \equiv (b, \tau, \alpha_0, \sigma^2)\). Before proceeding any further, observe that we can rewrite model~\eqref{eq:rkhs-model-linear} in a more explicit and practical fashion in terms of the available sample information in \(\mathcal D_n\). For \(\theta \in \Theta_p\), the reinterpreted model assumes the form
\begin{equation}\label{eq:rkhs-model-linear-2}
  Y_i \mid X_i, \theta \ \stackrel{\text{i.i.d.}}{\sim} \mathcal N\left(\alpha_0 + \sum_{j=1}^p \beta_j X_i(t_j), \ \sigma^2\right), \quad i =1,\dots, n.
\end{equation}

It is worth mentioning that the model remains linear in the sense that it fundamentally involves a random variable belonging to the linear span of the process \(X\) in \(L^2(\Omega)\). Also, note that given the time instants \(t_j\) the model becomes a multiple linear model with the \(X(t_j)\) as scalar covariates. As a matter of fact, this RKHS model is particularly suited as a basis for variable selection methods, and also entails the classical \(L^2\)-model~\eqref{eq:l2-linear-model} under certain conditions \citep[see][Sec.~3]{berrendero2020general}. In addition, as far as generalizations go the proposed model is easily extensible, seeing as how several functional covariates could be added to obtain an extended model such as \(Y=\alpha_0 + \Psi_{X^{1}}(\alpha_1) + \cdots + \Psi_{X^{q}}(\alpha_q) + \epsilon\). In that case, as argued in \citet{grollemund2019bayesian} for a similar situation, if we were to set a prior distribution on all the parameters involved, we could recover the full posterior by looking alternately at the posterior distribution of each covariate conditional on the rest of them.

The prior distribution suggested for the parameter vector \(\theta \in \Theta_p\) is given by
\begin{align}\label{eq:prior-linear}
  \begin{split}
  \pi(\alpha_0, \sigma^2)              & \propto 1/\sigma^2,                                                     \\
  \tau                     & \sim \mathcal U([0, 1]^p),                                              \\
  b\mid \tau, \sigma^2 & \sim \mathcal N_p(b_0, g\sigma^2(G_\tau + \eta I)^{-1}),
\end{split}
\end{align}
where \(I\) is the identity matrix, \(G_\tau = \mathcal X_\tau'\mathcal X_\tau\) is the Gram matrix corresponding to the matrix \(\mathcal X_\tau = (X_i(t_j))_{i,j}\), and \(b_0\in \R^p, \ g \in \R\) and \(\eta \in \R^+\) are hyperparameters of the model. On the one hand, note the use of a joint prior distribution on \(\alpha_0\) and \(\sigma^2\), which is a widely used non-informative prior known as Jeffrey's prior \citep{jeffreys1946invariant}. In any event, the estimation of \(\alpha_0=\E[Y]\) is straightforward, so it could have been left out of the model altogether. On the other hand, the prior on \(b\) is a slight modification of the well-known Zellner's g-prior \citep{zellner1986assessing}, in which a regularizing term is added to avoid ill-conditioning problems in the Gram matrix, obtaining a ridge-like Zellner prior controlled by the tuning parameter \(\eta\) \citep{baragatti2012study}. All in all, the proposed prior distribution becomes \(\pi(\theta) = \pi(\alpha_0, \sigma^2)\pi(\tau)\pi(b\mid \tau, \sigma^2)\).

As for the posterior distribution, a standard algebraic manipulation in~\eqref{eq:bayes-theorem} yields the following result.

\begin{proposition}
Under model~\eqref{eq:rkhs-model-linear-2}, the prior distribution implied in~\eqref{eq:prior-linear} produces the posterior distribution
\[
\pi(\theta \mid \mathcal D_n) \propto \frac{|G_\tau|^{1/2}}{\sigma^{p+n}} \exp\left\{ -\frac{1}{2\sigma^2} \left(\|Y- \alpha_0\bm{1} - \mathcal X_\tau b\|^2 + \frac{1}{g}(b - b_0)'G_\tau(b - b_0) \right) \right\},
\]
where \(Y=(Y_1,\dots,Y_n)'\) and \(\bm{1}\) is an \(n\)-dimensional vector of ones.
\end{proposition}

Let us recall that when performing the empirical posterior approximation, on each of the \(M\) steps of the iterative MCMC algorithm we get an approximate sample \(\theta^{(m)*}\) of the posterior distribution \(\pi(\theta\mid \mathcal D_n)\). Assuming now a previously unseen test set \(\mathcal D'_{n'}\) in the same conditions as \(\mathcal D_n\), we propose to construct three different kinds of predictors based on the MCMC samples, each of them following a different strategy.

\begin{description}
  \item[Summarize-then-predict] If we consider a point-estimate statistic \(T\) to act as a summary of the posterior distribution, we can get the corresponding estimates \(\hat{\theta}=(\hat b, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T(\{\theta^{(m)*}\})\) and predict the responses in the usual way \incomment{Following model...} by
  \[
    \hat Y_i = \hat{\alpha}_0 + \sum_{j=1}^p \hat{\beta}_j X_i(\hat{t}_j), \quad i=1,\dots, n'.
  \]
  \incomment{Pasar a la sección 2.3} The choice of a specific summary statistic results in a veiled assumption of an underlying loss function between the estimated and real parameters. In general, the mean is more sensitive to outliers while the median is more robust, but the latter assumes a \(L^1\)-type loss function. On the other hand, the mode is also a good candidate because it represents the point of highest probability. At any rate, these decisions are strongly dependent on several factors such as the skewness or the number of modes in the resulting posterior distribution, and thus should be made on a case-by-case basis.
  \item[Predict-then-summarize] We can also look at the approximate posterior distribution as a whole and compute the distribution of the ``sampled responses'' at each step of the chain following model~\eqref{eq:rkhs-model-linear-2}:
  \[
  Y^{(m)*} := \left\{Y_i^{(m)*} \equiv Y_i \mid X_i, \theta^{(m)*}:\ i=1,\dots,n'\right\}, \quad m=1,\dots,M.
  \]
  Then, we can take the mean of all such samples as a proxy for each response variable:
  \[
  \hat Y_i = \frac{1}{M}\sum_{m=1}^M Y_i^{(m)*}, \quad i=1,\dots,n'.
  \]

  \begin{comment}
   Separar un poco lo de selección de variables, comentando que nos concentramos solo en los tiempos y descartamos el resto (se consideran nuisance) (promediar los subvectores de tiempos y luego predecir como en la estrategia A, pero usando modelo de regresión multiple).
  \end{comment}
  \item[Variable selection] Lastly, we can focus only on the marginal posterior distribution of \(\tau\mid\mathcal D_n\) and perform variable selection via a point-estimate statistic \(T\). Specifically, we can consider the times \(\hat t_j = T(\{t_j^{(m)*}\})\) and reduce the original data set to the scalar variables \(\{X_i(\hat t_j):\ i=1, \dots,n, \ j=1,\dots,p\}\). After this selection, we can treat the problem as a finite-dimensional regression task and use any of the well-known prediction algorithms suited for this situation.
\end{description}

@@

In the case of functional logistic regression, we can view the binary response variable \(Y\in\{0, 1\}\) as a Bernoulli random variable given the regressor \(X=x \in L^2[0, 1]\), and suppose that \(\log\left(p(x)/(1-p(x))\right)\) is linear in \(x\), where \(p(x)=\mathbb P(Y=1\mid X=x)\). Then, a RKHS approach given by the following functional logistic equation:
\begin{equation}\label{eq:rkhs-model-logistic}
  \mathbb P(Y=1 \mid X) = \frac{1}{1 + \exp\{-\alpha_0 - \Psi_X(\alpha)\}}, \quad \alpha_0 \in \R, \ \alpha \in \Hcal_{0,p}(K).
\end{equation}
Following our previous strategy, after incorporating the sample information we can rewrite the classification model~\eqref{eq:rkhs-model-logistic} as
\begin{equation}\label{eq:rkhs-model-logistic-2}
Y_i \mid X_i,\theta \ \stackrel{\text{i.i.d.}}{\sim} \operatorname{Bernoulli}(p_i), \quad i=1,\dots, n,
\end{equation}
with
\begin{equation}\label{eq:rkhs-model-logistic-2-parameter}
  p_i = \mathbb P(Y_i=1 \mid X_i,\theta) = \frac{1}{\displaystyle 1 + \exp\left\{-\alpha_0 - \sum_{j=1}^p \beta_j X_i(t_j)\right\}}, \quad i=1,\dots, n,
\end{equation}
where in turn \(\beta_j\in\R\), \(t_j\in[0, 1]\) and \(\alpha_0\in\R\).

@@

As far as prior distributions go, we propose to use the same ones as we did in~\eqref{eq:prior-linear} for the linear regression approach. However, in this case the parameter \(\sigma^2\) is a nuisance parameter and can be safely ignored at the end, since it does not appear in the final model~\eqref{eq:rkhs-model-logistic-2}.

\begin{comment}
  Misma pregunta que antes: ¿ponemos la expresión de la posteriori?
  \begin{align*}
\log \pi(\theta \mid \mathcal D_n) & \propto \sum_{i=1}^n \left[ \left(\alpha_0 + \Psi_{x_i}(\alpha)\right)y_i - \log\left(1 + \exp\left\{\alpha_0 + \Psi_{x_i}(\alpha)\right\}\right)\right]\\
\quad &+ \frac{1}{2}\log |G_\tau| - p\log \sigma -\frac{1}{2g\sigma^2} (b - b_0)'G_\tau(b - b_0).
\end{align*}

@@

Lastly, the variable selection method is essentially the same as in the case of linear regression: we select \(p\) times from each trajectory based on the posterior distribution of \(\tau\mid \mathcal D_n\), and then feed the reduced dataset to a finite-dimensional binary classification procedure.

@@

One of the key decisions in our Bayesian modeling scheme was whether to consider the number of components \(p\) as a member or the parameter space and integrate it into the model. While theoretically we could impose a prior distribution on \(p\) as well, we found that it would have some unwanted practical implications. For instance, it would make the implementation more complex, since the dimensionality of the parameters \(b\) and \(\tau\) would need to be fixed at a certain maximum value, but in the MCMC algorithm the working value of \(p\) would vary from one iteration to the next. In this case we would face a label switching issue \citep[c.f.][Sec.~2.3]{grollemund2019bayesian}, as we would have no way of knowing which set of parameters is ``active'' at any given time. Another possibility would be to adapt a purely Bayesian model selection technique \citep[see][Ch.~7]{piironen2017comparison, gelman1995bayesian} to our framework, or even derive some model aggregation methods to combine the posterior distributions obtained for different-sized models. In the end, for the sake of simplicity we decided to let \(p\) be an hyperparameter, so that we could use any model selection criteria (e.g. BIC, DIC, \ldots) to select its optimal value \incomment{deberíamos elegir uno de los criterios?}.
